\bf Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-03-16 doc/paper.Rmd
couple more citations

## Motivation
Same source analyses are a major part of an Forensic Toolmark Examiner's job. In current practice examiners  make these comparisons visually under a comparison microscope and come to one of the following four conclusions: identification, inconclusive, elimination or unsuitable for examination \citep{afte-toolmarks1998}. These conclusions are made on the basis of "unique surface contours" of the two toolmarks being in "sufficient agreement" \citep{afte-toolmarks1998}. AFTE describes the term "sufficient agreement" as the possibility of another tool producing the markings under comparison, as practically impossible \citep{afte-toolmarks1998}. This subjectivity in the assessment as well as the lack of error rates are the main points of criticisms in the toolmark examination first raised by the National Research Council in 2009 \citep{NAS:2009} and later emphasized further by the President's Council of Advisors on Science and Technology \citep{pcast2016}.

Technological advances, such as profilometers and confocal microscopy allow to measure 3D surfaces \citep{vorburger2016} in a digitized form. This forms the basis of statistical analysis of toolmarks. A statistical approach based on data  removes both subjectivity from the assessment and allows a quantification of error rates for both false positive and false negative identifications.

Various toolmarks have been studied in the literature:  screwdriver marks  digitized using a profilometer have been analyzed in \citet{manytoolmarks1} and \citet{chumbley};  \citet{manytoolmarks2} have investigated 3D marks from screwdriver, tongue and groove pliers captured using a confocal microscope;  digitized marks from slip-joint pliers generated by a surface profilometer have been investigated by \citet{afte-chumbley}.

\citet{manytoolmarks2} define a relative distance metric and use it as similarity measure between two toolmarks. \citet{manytoolmarks1} extract many small segments in the markings of two toolmarks and compare similiarity using a maximum pearson correlation coefficient. 
The Chumbley scoring method, first introduced by \citet{chumbley}, uses a similar but more extensive framework based on a Mann-Whitney U test of the resulting correlation coefficients. This approach was later improved by Hadler and Morris \citep{hadler} where they introduce a deterministic approach and address the issue of lack of independece between segements of striae.  In this paper, we are investigating the applicability of the Chumbley scoring method by \citet{hadler} to assess striation marks on bullet lands for same-source identification.

Striation marks on bullets are made by impurities in the barrel. As the bullet travels through the barrel, these imperfections leave "scratches" on the bullet surface. Typically, only striation marks in the land engraved areas (LEAs) are considered \citet{afte-article1992}. Bullet lands are depressed areas between the  grooves made by the rifling action of the barrel. Compared to toolmarks made by screwdrivers striation marks on bullets are typically much smaller, both in length and in width. Bullets also have a curved cross-sectional topography \hh{sketch?}. Figure \ref{fig:rgl} shows us how the signature from a bullet land lines up with the image of the land from which it was extracted. We can also see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as the signature.

Bullet matching methods depend on the attribute being used for making comparisons. For single feature based matching \citet{chu2013} used an automatic method for counting consequtive matching striae (CMS). They reported an error rate 52% of the known same source lands comparisons as misidentified and zero false positives for known different source lands. \citet{ma2004} and \citet{vorburger2011} discuss about CCF (cross-correlation function) and its discriminating power and application but do not talk about matches and non-matches. \citet{aoas} use multiple attributes like CCF, CMS, D (distance measure) etc in a random forrest based method and compare every land against every other land of the Hamby dataset \citet{hamby}. They report an out-of-bag error rate of 0.0046.

The Chumbley score  provides us with another approach in the same-source assessment of bullet striation marks.\citet{chumbley}, in their paper, compare two toolmarks with the intention of determining if it comes from the same tool. The data for their study was obtained from 50 sequentially manufactured screwdriver tips.\citet{chumbley} reported error rates for marking made by the tips at differnt angles. For a 30 degree they report false negative error rate to be 0.023 and  false positive error rate be 0.09. For other angles the error rates remain the same for false negatives and 0.01 for false positives. The improvement proposed by \citet{hadler} on the other hand, did not consider effect of angles and gave 0.06 false  negative error rate and 0 false poisitive error rate for markings made at the same angle.

>>>>

## Motivation
Same source analyses are a major part of an Forensic Toolmark Examiner's job. In current practice examiners  make these comparisons by visual inspection under a comparison microscope and come to one of the following four conclusions: identification, inconclusive, elimination or unsuitable for examination~\citep{afte-toolmarks1998}. These conclusions are made on the basis of "unique surface contours" of the two toolmarks being in "sufficient agreement" \citep{afte-toolmarks1998}. AFTE describes the term "sufficient agreement" as the possibility of another tool producing the markings under comparison, as practically impossible \citep{afte-toolmarks1998}. This subjectivity in the assessment as well as the lack of error rates are the main points of criticisms  first raised by the National Research Council in 2009 \citep{NAS:2009} and later emphasized further by the President's Council of Advisors on Science and Technology \citep{pcast2016}.

Technological advances, such as profilometers and confocal microscopy allow to measure 3D surfaces in a high-resolution digitized form. This technology has become more accessible over the last decade, and has made its way into topological images of ballistics evidence, such as bullet lands and breech faces \citep{DeKinder1, DeKinder2, Bachrach1, vorburger2016}.
Digitized images of 3D surfaces of form the data basis of statistical analysis of toolmarks. A statistical approach based on data  removes both subjectivity from the assessment and allows a quantification of error rates for both false positive and false negative identifications.

\hh{In the next page and a half it is easy to lose the red line. It might help to include a table with an overview.
The table should include the reference to the paper, the data used, the statistical method and the associated error rates.}
Various toolmarks have been studied in the literature:  \citet{manytoolmarks1} and \citet{chumbley} have been analyzing screwdriver marks  digitized using a profilometer;  \citet{manytoolmarks2} have investigated 3D marks from screwdriver, tongue and groove pliers captured using a confocal microscope; \citet{afte-chumbley} have been investigated digitized marks from slip-joint pliers generated by a surface profilometer.

\hh{We need an additional sentence here to get from the data to the statistical methods ... }
\citet{manytoolmarks2} define a relative distance metric and use it as similarity measure between two toolmarks. \citet{manytoolmarks1} extract many small segments in the markings of two toolmarks and compare similiarity using a maximum pearson correlation coefficient. 
The Chumbley scoring method, first introduced by \citet{chumbley}, uses a similar but more extensive framework based on a Mann-Whitney U test of the resulting correlation coefficients. This approach was later improved by Hadler and Morris \citep{hadler} where the authors introduce a deterministic approach and address the issue of lack of independence between segements of striae.  In this paper, we are investigating the applicability of the Chumbley scoring method by \citet{hadler} to assess striation marks on bullet lands for same-source identification.

Striation marks on bullets are made by impurities in the barrel. As the bullet travels through the barrel, these imperfections leave "scratches" on the bullet surface. Typically, only striation marks in the land engraved areas (LEAs) are considered \citet{afte-article1992}. Bullet lands are depressed areas between the  grooves made by the rifling action of the barrel. Compared to toolmarks made by screwdrivers striation marks on bullets are typically much smaller, both in length and in width. Bullets also have a curved cross-sectional topography. Figure~\ref{fig:rgl} shows us how the signature from a bullet land (bottom) lines up with the image of the land (top) from which it was extracted. We can also see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as the signature.

\hh{Here's a bit of a jump: you're already talking about features, but we haven't yet properly introduced signatures of bullets yet.}
Bullet matching methods depend on the attributes  used for making comparisons. For single feature based matching \citet{chu2013} used an automatic method for counting consecutive matching striae (CMS). The authors report an error rate of 52% of the known same source lands comparisons as misidentified (false negative) and zero false positives for known different source lands. \citet{ma2004} and \citet{vorburger2011} discuss CCF (cross-correlation function) and its discriminating power and applicability for same-source analyses of bullets, but do not provide any error rates in their discussion. \citet{aoas} use multiple features like CCF, CMS, D (distance measure) etc in a random forest based method and compare every land against every other land of  digitised versions of Hamby 252 and Hamby 44 \citep{hamby} published on the NIST Ballistics Database \citep{nist}. The authors report an out-of-bag overall error rate of 0.46%, comprised of a false positive error rate of 30.05% and a false negative rate of 0.026%.

The Chumbley score  provides us with another approach in the same-source assessment of bullet striation marks. \citet{chumbley}, in their paper, compare two toolmarks with the intention of determining if it comes from the same tool. The data for their study was obtained from 50 sequentially manufactured screwdriver tips. \citet{chumbley} reported error rates for marking made by the tips at differnt angles. For a 30 degree they report false negative error rate to be 0.023 and  false positive error rate be 0.09. For other angles the error rates remain the same for false negatives and 0.01 for false positives. The improvement proposed by \citet{hadler} on the other hand, did not consider effect of angles and gave 0.06 false  negative error rate and 0 false poisitive error rate for markings made at the same angle.

### 2018-03-16 doc/paper.tex
couple more citations

\hh{XXX first sentence: Same source analyses are a major part of an Forensic Toolmark examiners job. XXX second sentence: describe standard practice: In current practice firearm examiners  make these comparisons visually under a comparison microscope and report XXX what is being reported: AFTE. XXX find: AFTE theory of identification as it relates to toolmarks  }
\hh{One of the main criticisms in this toolmark examination first raised by the National Research Council in 2009 [@NAS:2009] and later emphasized further by the President's Council of Advisors on Science and Technology [@pcast2016] are the amount of subjectivity in the assessment and the lack of error rates. }
\hh{XXX we have digitized images, e.g. profilometers or Confocal microscopy allow us to measure the surface}

\hh{Using a statistical approach removes both subjectivity from the assessment and allows us to provide error rates for both false positive and false negative identifications. }
\hh{Multiple toolmarks have been studied in the literature:  \citep{manytoolmarks1, manytoolmarks2, chumbley} have studied screwdriver marks digitized using a profilometer,  \citep{manytoolmarks2} investigate groove pliers XXX data? XXX  to slip-joint pliers \citep{afte-chumbley} XXX. In all of these cases, the statistical test used, is a variation of the Chumbley scoring method, first introduced by XXX and later improved by XXX. }

\hh{XXX In this pper, we are investigating the applicability of the Chumbley scoring method by @hadler to assessing striation marks on bullets for same-source identification. }

\hh{Striation marks on bullets are made by impurities in the barrel. As the bullet travels through the barrel, these imperfections leave "scratches' on the bullet surface. Typically, only striation marks in the land engraved areas (LEAs) are considered (XXX citation) [@afte-article1992]. Compared to toolmarks made by screwdrivers striation marks on bullets are typically much smaller, both in length and in width. Bullets also have a curved cross-sectional topography. 
}

Compairing pairs of toolmarks with the intention of matching it to a
tool has been studied many times in the past. Extensive examples can be
found in literature for tools and toolmark research ranging from
screwdrivers \citep{manytoolmarks1, manytoolmarks2, chumbley} to groove
pliers \citep{manytoolmarks2} to slip-joint pliers \citep{afte-chumbley}
and many more. In comparison to this, same source matching of bullets to
firearms has not been examined as prominently as that of toolmarks. Even
less information is available on validity of methods and error rates
associated with firearms examination. The National Academy of Sciences
in its report in 2009 \citep{NAS:2009} discussed the need for
determining error rates in methods proposed for firearms examination. As
seen in the case of most forensic applications, the first step for same
source matching and error rate determination involves identification of
unique features that are characteristic of the object at hand. For the
case of bullets and firearms, striation marks on the surface of the
bullet are considered to be such markings that can be used in methods
for same source matching. These marks are often a product of rifling and
impurities and defects due to manufacturing in the barrel of the gun,
which leads to engravings on the bullet surface
\citep{afte-article1992}. In current practice, firearm examiners
invariably make visual comparisons of bullet striae and use visual
assesment tools to dignify bullets as being matches and non-matches. One
way of accomplishing anykind of comparison between bullets is to do a
comparison between surface marking of two or more bullet lands. Bullet
Lands are considered to be areas between grooves made by the rifling
action of the barrel and the markings on them are considered to be
unique. The land engraved markings or sometimes termed as Bullet
profiles \citep{aoas} are striation marks made on Bullet lands and often
used for these land to land comparisons. Bullet Signatures is another
word used in literature as seen in the work of \citet{chu2013} and
\citet{aoas}. In our context bullet signatures refer to a processed
version of the raw land engraved markings or profiles. The generation of
bullet signatures involves first extraction of a bullet profile by
taking the cross-sectional of the surface at a given height and then
using loess fits to model the structure. The residuals of this fit are
called signatures, which are considered to be noise free and a good
reflection of the class charecteristics and unique features of a bullet.
A more detailed version of the extraction technique of signatures is
discussed by \citet{aoas}, where comprehensive details about the height
at which profile is to be selected, removing curvature, smoothing,
identifying groove locations are explained. Figure \ref{fig:rgl} shows
us how the extracted signature from a bullet land lines up with the
image of the land from which it was extracted. We can see also see in
the figure how the depth and relative position of the striation markings
seen in the image are interpreted as the signature.

As mentioned earlier, subject bias and error rate determination have
been a long standing issue in firearm examination \citep{NAS:2009}. This
issue of removing subjectivity and making firearm analysis objective was
also raised in the \citet{pcast2016} report. For an objective analysis,
scientific principle testability and error rate determination are
considered to be defining aspects. Identification of the uncertainities
associated with a method of comparison and coming up with a systemized
method of quantification, is therefore important. Associating
mathematical probabilites in a systematic manner to these methods are
one such way of quantification. In a previous study conducted by
\citet{aoas} a machine learning based algorithm was developed for same
source matching of bullets and error rates were discussed using the
database from the Hamby Study \citep{hamby}. The method proposed by
\citet{chumbley} (and later improved by \citet{hadler}) also provides a
means to determine error rates and claims to reduce subject bias.
Therefore, giving us a strong motivation to explore the adaptibility of
the Chumbley score methodology to bullets. In this paper, we take this
toolmark based methodology and try to apply and adapt it to bullets.
After this, we consequently discuss about the efforts in doing so, along
with the associated error rates. The data used in this paper also
belongs to the Hamby Study \citep{hamby}. This gives us a common
platform for comparing the performance of the chumbley method on bullets
with the already existing method proposed by \citet{aoas} for bullets.
As a brief overview, \citet{chumbley}, in their paper, compare two
toolmarks with the intention of determining if it comes from the same
tool. They use an empirical based setup to validate their proposed
algorithm and quantitative method which calculates a U-statisic for the
purpose of classification of toolmarks as matching or non-matching. The
data for their study was obtained from 50 sequentially manufactured
screwdriver tips, and preselected comparison window sizes were given as
inputs to the algorithm. The algorithm then compares the two toolmarks
and comes up with a Mann-Whitney U-statistic and an associated p-value
to designate them as matches or non-matches. The performance for every
100 comparisons, of the algorithm proposed by \citet{chumbley} and the
improvement proposed by \citet{hadler} are listed in the table below.

\begin{table}[!htb]
    \begin{minipage}{.5\linewidth}
      \caption{Chumbley et al. 2010}
      \centering 
\begin{tabular}{lrr}
\toprule
Classification & Match & Non-Match\\
\midrule
Match & 41 & 9\\
Non-Match & 2 & 48\\
\bottomrule
\end{tabular} \end{minipage}%
    \begin{minipage}{.5\linewidth}
      \centering
        \caption{Hadler et. al. (2017)} 
\begin{tabular}{lrr}
\toprule
Classification & Match & Non-Match\\
\midrule
Match & 47 & 3\\
Non-Match & 0 & 50\\
\bottomrule
\end{tabular} \end{minipage} 
\end{table}

\subsection{Potential limitations of the Chumbley Score adaptation to
bullets}\label{potential-limitations-of-the-chumbley-score-adaptation-to-bullets}

Bullet striations are different from toolmarks. Bullet markings are
typically produced by the barrel on the bullets, while toolmarks can be
made by tool tips on different surfaces. Tools like screw driver tips as
used by \citet{chumbley}, produce longer and pronounced markings.
Bullets on the other hand are much smaller in length, width, and curved
in the cross-sectional topography. This means the markings made by
barrels as opposed to markings made by tools, may have a problem in
distinctiveness. The majority of Bullet profiles and signatures of the
Hamby study \citep{hamby} used in this paper are extracted by procedures
mentioned by \citet{aoas}. The Ruger pistol barrels used in the Hamby
study have a 9mm caliber. This means that bullets used in these pistols
have land markings of almost \((1/6)^{th}\) the size of screwdriver
toolmarks used by \citet{chumbley}. Also, striations on bullets are made
on their curved surfaces, whereas the algorithm developed by
\citet{chumbley} and \citet{hadler} has only been tested for flatter and
wider surfaces which have negligible curvature. Therefore, using methods
proposed for toolmarks may need adaptation in order to give tangible
results for bullets. Moreover, in order to get flat bullet signatures
and remove the curvatures some kind of smoothing needs to be applied as
a pre-step wihch needs further investigation as to whether the level of
smoothing does effect the working of the algorithm on Bullets.

The Chumbley method and algorithm works on the premise of comparison
windows which are supposed to represent small segments of pixels of
equal sizes in the two markings that are to be compared in a
predetermined fashion. They term them as windows of optimization and
validation. The optimization window is selected in the first step called
the optimization step where areas/windows of best agreement between the
the two markings are found. The window of optimization for bullet
striations are shorter as bullet signatures are smaller compared to
toolmarks . The idea is to keep the number of windows of optimization
sufficiently large. This means we have shorter trace segments that lets
us compare smaller segments of one signature to another than those for
toolmarks. Here, trace segments are the partitions of marking with the
length of the partition equal to the size of window of optimization.
This introduces a problem as, if we go too small in the window of
optimization, the unique features of the trace segments are lost and
seem similar, while too large sizes vastly reduces the weight of small
features that would otherwise uniquely classify a signature and hence
identify the region of agreement. The problem of distinctiveness in the
structure of bullet striations may also lead to potential failure of the
algorithm to identify correct windows of agreements or not be able to
identify these windows at all. This may lead to incorrect classification
of bullets as matches or non-matches.

@@

@@

The Chumbley score algorithm takes input as two vectorized processes.
Here the processes are of the form \(z(t)\) which is a spatial process
for some location indexed with \(t\). This means that while \(z(t)\)
represents any striae, \(z(t_1)\) is a realization of the spatial
process. \(t\) here denotes a vector of equally spaced pixel locations
for the striation marks under consideration. The word `pixel' refers to
the resolution of the confocal light microscope. In the case for bullet
signatures extracted from NIST ballistics database \citep{nist} of the
Hamby study, a pixel corresponds to approximately 1.56 microns. The
vectorized processes representing two sets of striaes are therefore
shown as \(x(t_1)\), \(t_1 = 1,2,...T_1\) and \(y(t_2)\),
\(t_2 = 1,2...T_2\). Here \(x\) and \(y\) denote two instances of the
process \(z(*)\) which means two striation marks, while the indexing
\(t\) or pixel range is shown as \(t_1\) and \(t_2\) for \(x\) and \(y\)
respectively. This representation of \(t\) as \(t_1\) and \(t_2\) lets
us identify them as either of same or different lengths. The striation
marks under consideration are potentially from two different bullets
whose source needs to be identified as being same or different. \(T_1\)
and \(T_2\), as represented above, are the final pixel indexes of each
marking and therefore give the respective lengths of the markings. The
similarity in the two markings is then judged by the algorithm on the
basis of cross-correlation of a fixed and constant number consecutive
pixels (say \(k\)) taken from the two markings. This can for example be
\(k\) taken from the two indexed marking \(x(t_1)\) and \(y(t_2)\) such
that in theory \(k\) remains smaller than length of the two striae or
marks. Depending on what stage of the algorithm we are in, matching of
different pixel lengths and locations is done. This in the end,
effectively compares all possible windows that would guarantee in
quantifying the two marks or striae as coming from the same source or
not.

@@

>>>>

Same source analyses are a major part of an Forensic Toolmark Examiner's
job. In current practice examiners make these comparisons by visual
inspection under a comparison microscope and come to one of the
following four conclusions: identification, inconclusive, elimination or
unsuitable for examination\textasciitilde{}\citep{afte-toolmarks1998}.
These conclusions are made on the basis of ``unique surface contours''
of the two toolmarks being in ``sufficient agreement''
\citep{afte-toolmarks1998}. AFTE describes the term ``sufficient
agreement'' as the possibility of another tool producing the markings
under comparison, as practically impossible \citep{afte-toolmarks1998}.
This subjectivity in the assessment as well as the lack of error rates
are the main points of criticisms first raised by the National Research
Council in 2009 \citep{NAS:2009} and later emphasized further by the
President's Council of Advisors on Science and Technology
\citep{pcast2016}.

Technological advances, such as profilometers and confocal microscopy
allow to measure 3D surfaces in a high-resolution digitized form. This
technology has become more accessible over the last decade, and has made
its way into topological images of ballistics evidence, such as bullet
lands and breech faces
\citep{DeKinder1, DeKinder2, Bachrach1, vorburger2016}. Digitized images
of 3D surfaces of form the data basis of statistical analysis of
toolmarks. A statistical approach based on data removes both
subjectivity from the assessment and allows a quantification of error
rates for both false positive and false negative identifications.

Various toolmarks have been studied in the literature:
\citet{manytoolmarks1} and \citet{chumbley} have been analyzing
screwdriver marks digitized using a profilometer; \citet{manytoolmarks2}
have investigated 3D marks from screwdriver, tongue and groove pliers
captured using a confocal microscope; \citet{afte-chumbley} have been
investigated digitized marks from slip-joint pliers generated by a
surface profilometer.

\hh{We need an additional sentence here ... } \citet{manytoolmarks2}
define a relative distance metric and use it as similarity measure
between two toolmarks. \citet{manytoolmarks1} extract many small
segments in the markings of two toolmarks and compare similiarity using
a maximum pearson correlation coefficient. The Chumbley scoring method,
first introduced by \citet{chumbley}, uses a similar but more extensive
framework based on a Mann-Whitney U test of the resulting correlation
coefficients. This approach was later improved by Hadler and Morris
\citep{hadler} where the authors introduce a deterministic approach and
address the issue of lack of independence between segements of striae.
In this paper, we are investigating the applicability of the Chumbley
scoring method by \citet{hadler} to assess striation marks on bullet
lands for same-source identification.

Striation marks on bullets are made by impurities in the barrel. As the
bullet travels through the barrel, these imperfections leave
``scratches'' on the bullet surface. Typically, only striation marks in
the land engraved areas (LEAs) are considered \citet{afte-article1992}.
Bullet lands are depressed areas between the grooves made by the rifling
action of the barrel. Compared to toolmarks made by screwdrivers
striation marks on bullets are typically much smaller, both in length
and in width. Bullets also have a curved cross-sectional topography.
Figure\textasciitilde{}\ref{fig:rgl} shows us how the signature from a
bullet land (bottom) lines up with the image of the land (top) from
which it was extracted. We can also see in the figure how the depth and
relative position of the striation markings seen in the image are
interpreted as the signature.

\hh{Here's a bit of a jump: you're already talking about features, but we haven't yet properly introduced signatures of bullets yet.}
Bullet matching methods depend on the attributes used for making
comparisons. For single feature based matching \citet{chu2013} used an
automatic method for counting consecutive matching striae (CMS). The
authors report an error rate of 52\% of the known same source lands
comparisons as misidentified (false negative) and zero false positives
for known different source lands. \citet{ma2004} and
\citet{vorburger2011} discuss CCF (cross-correlation function) and its
discriminating power and applicability for same-source analyses of
bullets, but do not provide any error rates in their discussion.
\citet{aoas} use multiple features like CCF, CMS, D (distance measure)
etc in a random forest based method and compare every land against every
other land of digitised versions of Hamby 252 and Hamby 44 \citet{hamby}
published on the NIST Ballistics Database \citep{nist}. The authors
report an out-of-bag overall error rate of 0.46\%, comprised of a false
positive error rate of 30.05\% and a false negative rate of 0.026\%.

The Chumbley score provides us with another approach in the same-source
assessment of bullet striation marks.\citet{chumbley}, in their paper,
compare two toolmarks with the intention of determining if it comes from
the same tool. The data for their study was obtained from 50
sequentially manufactured screwdriver tips.\citet{chumbley} reported
error rates for marking made by the tips at differnt angles. For a 30
degree they report false negative error rate to be 0.023 and false
positive error rate be 0.09. For other angles the error rates remain the
same for false negatives and 0.01 for false positives. The improvement
proposed by \citet{hadler} on the other hand, did not consider effect of
angles and gave 0.06 false negative error rate and 0 false poisitive
error rate for markings made at the same angle.

@@

\subsection{Scans for land engraved
areas}\label{scans-for-land-engraved-areas}

The land engraved markings across the cross section of the land are
termed as bullet profiles \citep{aoas,ma2004}. These are often used for
making land to land comparisons. Bullet signatures on the other hand
\citep{chu2013,aoas} refer to a processed version of the raw land
engraved markings or profiles. The generation of bullet signatures
involves first extraction of a bullet profile by taking the
cross-sectional of the surface at a given height. Loess fits are then
used to model the structure. The residuals of these fits are called
signatures. They are considered to be noise free and a good reflection
of the key attributes of the raw marking, and the unique features of a
bullet. A more detailed version of the extraction technique of
signatures is discussed by \citet{aoas}, where comprehensive details
about the height at which profile is to be selected, removing curvature,
smoothing, identifying groove locations are explained.

The majority of Bullet profiles and signatures of the Hamby study
\citep{hamby} used in this paper are extracted by procedures mentioned
by \citet{aoas}. The scans for the Hamby study come from two sources
namely, NIST scans \citep{nist} and scans produced at Iowa State
University. Both scans were generated using confocal microscopes but at
different resolutions. The NIST database contains scans of the bullet
sets Hamby 44 and 252 with the resolution 1.5625 \(\mu m\) per pixel.
The Iowa State scans are available for only the Hamby 44 bullet set and
at a resoultion of 0.65 \(\mu m\) per pixel. The data used by
\citet{chumbley, hadler} on the other hand, was produced by a
profilometer at a resolution of about 0.73 \(\mu m\) per pixel. The
different resolutions indicate that the length of the digitized markings
would be significantly different for each setting.

\subsection{Potential Challenges in Chumbley Score
Adaptation}\label{potential-challenges-in-chumbley-score-adaptation}

Markings made by Screw driver tips \citep{chumbley} are longer and
pronounced and about 7 mm in length \citep{manytoolmarks1}, while the
groove to groove length of the bullet lands from the Hamby study
\citep{hamby} are about 2mm in length. The Ruger pistol barrels used in
the Hamby study have a 9mm caliber. This means that bullets used in
these pistols have land markings of almost \((1/4)^{th}\) the size of
screwdriver toolmarks used by \citet{chumbley, hadler}, therefore
raising a problem of distinctiveness. One way to investigate this is to
evaluate bullet striations at different resolutions. This gives us an
incentive to use both the NIST and Iowa State scans and investigate the
effect of length of the digitized mark on the working of the algorithm.
Apart from this, the algorithm developed by \citet{hadler} has only been
tested for flatter and wider surfaces which have negligible curvature,
whereas striations on bullets are made on their curved surfaces.
Therefore, using methods proposed for toolmarks need adaptation in order
to give tangible results for bullets. Another challenge that arises from
these topographical differences, is the understanding of that effect of
different levels of smoothing. Especially investigating its effect on
the algorithm performance when signatures are being compared and when
raw markings or profiles are being compared.

Some other challenges associated with the adaptation process call for a
brief understanding of the methodology involved. The Chumbley method and
algorithm works on the premise of comparison windows that are supposed
to represent small segments of pixels of equal sizes in the two
markings. These windows are to be compared in a predetermined fashion
and are termed as windows of optimization and validation. The
optimization window is selected in the first step called the
optimization step where areas/windows of best agreement between the the
two markings are found.

The window of optimization for bullet striations are shorter as bullet
signatures in terms of absolute length (in mm), are smaller compared to
toolmarks. Even in terms of digitized pixels, there is no
straightforward way of establishing what window size is to be chosen.
One important way to address this is to keep the number of windows of
optimization sufficiently large or not too small. This means, if the
number of windows are too small we inadvertently use bigger window sizes
as we want to compare all segements of one mark with the segments of the
other mark. This leads to not giving enough emphasis to smaller
segements of the two marking that might be in high agreement. In terms
of absolute length we end up comparing smaller segments between the two
markings while dealing with bullet lands, than those being compared for
toolmarks.

For too large window sizes, the weight of small features that would
otherwise uniquely classify a signature and hence identify the region of
agreement is vastly reduced. On the other hand, this introduces another
problem as, if we go too small in the window of optimization, the unique
features of the trace segments are lost and all seem similar. The
problem of distinctiveness in the structure of bullet striations may
also lead to potential failure of the algorithm to identify correct
windows of agreements or not be able to identify these windows at all.
This may lead to incorrect classification of bullets as coming from the
same-source (or matches) or from different source (or non-matches).

@@

The Chumbley score algorithm takes input as two vectorized processes.
Here the processes are of the form \(z(t)\) which is a spatial process
for some location indexed with \(t\). This means that while \(z(t)\)
represents any striae, \(z(t_1)\) is a realization of the spatial
process. \(t\) here denotes a vector of equally spaced pixel locations
for the striation marks under consideration. The vectorized processes
representing two sets of striaes are therefore shown as \(x(t_1)\),
\(t_1 = 1,2,...T_1\) and \(y(t_2)\), \(t_2 = 1,2...T_2\). Here \(x\) and
\(y\) denote two instances of the process \(z(*)\) which means two
striation marks, while the indexing \(t\) or pixel range is shown as
\(t_1\) and \(t_2\) for \(x\) and \(y\) respectively. This
representation of \(t\) as \(t_1\) and \(t_2\) lets us identify them as
either of same or different lengths. The striation marks under
consideration are potentially from two different bullets whose source
needs to be identified as being same or different. \(T_1\) and \(T_2\),
as represented above, are the final pixel indexes of each marking and
therefore give the respective lengths of the markings. The similarity in
the two markings is then judged by the algorithm on the basis of
cross-correlation of a fixed and constant number consecutive pixels (say
\(k\)) taken from the two markings. This can for example be \(k\) taken
from the two indexed marking \(x(t_1)\) and \(y(t_2)\) such that in
theory \(k\) remains smaller than length of the two striae or marks.
Depending on what stage of the algorithm we are in, matching of
different pixel lengths and locations is done. This in the end,
effectively compares all possible windows that would guarantee in
quantifying the two marks or striae as coming from the same source or
not.

@@

\paragraph{endedit}\label{endedit}

\hh{end of intro: remaining paper is structured as follows: introduce to the data we get from confocal microscopy, introduce to profiles and signatures.
Introduce to chumbley score method, apply chumbley and discuss results ....
}

\subsection{Scans for land engraved
areas}\label{scans-for-land-engraved-areas-1}

\begin{itemize}
\tightlist
\item
  scans available: NIST database (citation), Hamby 44 and Hamby 252
  (Hamby citation)
\item
  move figure up
\item
  discuss cross section, profiles and signatures
\end{itemize}
